---
title: "Linear Models"
author: "Alex Di Genova"
output: rmarkdown::github_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Linear models
### simple
```{r lm}
# Load necessary libraries
library(ggplot2)

# Load the data
data(mtcars)
head(mtcars)

# Fit a simple linear regression model
model <- lm(mpg ~ wt, data=mtcars)

# Summary of the model
summary(model)

# Plotting the data and the model
ggplot(mtcars, aes(x=wt, y=mpg)) +
  geom_point() +
  geom_smooth(method="lm", col="blue")
```
### multiple
```{r lmm}
# Fit a multiple linear regression model
model_mult <- lm(mpg ~ wt + hp + qsec, data=mtcars)

# Summary of the model
summary(model_mult)

# Diagnostic plots
par(mfrow=c(2,2))
plot(model_mult)

```
```{r}
# Coefficients
coef(model_mult)
# R-squared
summary(model_mult)$r.squared
# P-values
summary(model_mult)$coefficients[,4]
```

## Model Diagnosis

```{r }
plot(model_mult, which=1)  # Residuals vs Fitted plot
# Look for a random scatter of residuals around the horizontal line at zero.
```
```{r}
qqnorm(resid(model_mult))
qqline(resid(model_mult))
```
```{r}
plot(model_mult, which=3)  # Scale-Location plot
```
```{r}
library(car)
durbinWatsonTest(model_mult)
```

# Examples

1. Predicting Housing Prices
Objective: To predict the price of houses based on various features like the number of bedrooms, square footage, age of the house, and location.

```{r}
# Load necessary library
library(MASS)

# Load the Boston housing dataset
data("Boston")

# Fit a multiple linear regression model
#dis=weighted mean of distances to five Boston employment centres
#lstat=lower status of the population 
#medv = median value of owner-occupied homes in
#rm = average number of rooms per dwelling.
# age = proportion of owner-occupied units built prior to 1940.
housing_model <- lm(medv ~ lstat + rm + age + dis, data=Boston)

# Summary of the model
summary(housing_model)
```




2. Predicting Employee Salary

Objective: To predict employee salary based on years of experience, education level, and job role.
Dataset: A hypothetical dataset containing employee information.

```{r}
# Hypothetical dataset
employee_data <- data.frame(
  salary = c(60000, 70000, 80000, 120000, 100000, 95000, 75000),
  experience = c(5, 6, 7, 10, 8, 9, 6),
  education = c(1, 2, 2, 3, 2, 3, 2),  # 1: Bachelor's, 2: Master's, 3: PhD
  role = as.factor(c("Analyst", "Analyst", "Manager", "Manager", "Analyst", "Manager", "Analyst"))
)

# Fit a multiple linear regression model
salary_model <- lm(salary ~ experience + education + role, data=employee_data)

# Summary of the model
summary(salary_model)

```

experience: Number of years of experience (positive coefficient implies more experience leads to higher salary).
education: Education level (positive coefficient implies higher education levels lead to higher salary).
role: Job role (categorical variable, with different coefficients for different roles indicating their impact on salary).


3. Predicting Disease Progression
Objective: To predict the progression of a disease based on various biomarkers and patient characteristics.

Dataset: A hypothetical dataset containing patient information and biomarker levels.

```{r}
# Hypothetical dataset
disease_data <- data.frame(
  progression = c(3.2, 4.5, 2.8, 6.7, 5.1, 4.3, 3.9),
  biomarker1 = c(2.1, 3.4, 2.0, 5.0, 4.2, 3.8, 3.1),
  biomarker2 = c(1.5, 2.2, 1.7, 3.5, 2.8, 2.4, 2.0),
  age = c(45, 50, 39, 60, 55, 52, 47),
  gender = as.factor(c("M", "F", "M", "F", "M", "F", "M"))
)

# Fit a multiple linear regression model
disease_model <- lm(progression ~ biomarker1 + biomarker2 + age + gender, data=disease_data)

# Summary of the model
summary(disease_model)

```


## Todo

- Try to build a better model for the mtcars and housing dataset.
- Investigate how to select features to build the linear models.
- Explore predicting other vairables for both models.
