---
title: "TrioRF"
author: "Alex Di Genova"
date: "2022-09-26"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data

we load the matrix generated with the python code

```{r loadd}
r=read.table("kidONT.R94.kmers.txt",h=T)
head(r)
```

## Real clases

we do add the real class to the data

```{r addclases}
library(tidyverse)
r=r %>% mutate(class=if_else(startsWith(id,"K1"),"B","A")) 
```

## Baseline classification tool

This is how currently is performed the classification of paternal/maternal haplotypes which is based on the hits for each mer.

```{doc}
def compute_scaling_factors():
    num_kmers_a = kmers.get_number_kmers_in_set(haplotype_a_kmers)
    num_kmers_b = kmers.get_number_kmers_in_set(haplotype_b_kmers)
    max_num_kmers = max(num_kmers_a, num_kmers_b)
    scaling_factor_a = 1.0 * max_num_kmers / num_kmers_a
    scaling_factor_b = 1.0 * max_num_kmers / num_kmers_b
    return scaling_factor_a, scaling_factor_b
    
# determine the class
  hap_a_score = hap_a_count * scaling_factor_a
  hap_b_score = hap_b_count * scaling_factor_b

   if hap_a_score > hap_b_score:
     read_bin = "A"
  
  elif hap_b_score > hap_a_score:
        read_bin = "B"
   else:
    read_bin = "U"
   
```

we do implement the same strategy for R code
```{r baseline}
library(caret)

nk_hapa=c(1348958,1548667,1674112,1791266)
nk_hapb=c(963463,1138714,1262662,1376492)
sf_a=c(1,1,1,1)
sf_b=nk_hapa/nk_hapb
#k_15
r_k15=r %>% select(id,ha_k15,hb_k15,class) %>% 
  mutate(scoreA=ha_k15*sf_a[1],scoreB=hb_k15*sf_b[1], classp=if_else(scoreA > scoreB, "A",if_else(scoreA == scoreB, "U","B")))
r_k15[nrow(r_k15) + 1,] = c("dummy",0,0,"U",0,0,"U")
mk15=confusionMatrix(factor(r_k15$classp),factor(r_k15$class))
mk15

r_k18=r %>% select(id,ha_k18,hb_k18,class) %>% 
  mutate(scoreA=ha_k18*sf_a[2],scoreB=hb_k18*sf_b[2], classp=if_else(scoreA > scoreB, "A",if_else(scoreA == scoreB, "U","B")))
#we add a dummy class for classification purposes
r_k18[nrow(r_k18) + 1,] = c("dummy",0,0,"U",0,0,"U")
mk18=confusionMatrix(factor(r_k18$classp),factor(r_k18$class))
mk18

#k21 is used in trio-binnig

r_k21=r %>% select(id,ha_k21,hb_k21,class) %>% 
  mutate(scoreA=ha_k21*sf_a[3],scoreB=hb_k21*sf_b[3], classp=if_else(scoreA > scoreB, "A",if_else(scoreA == scoreB, "U","B")))
#we add a dummy class for classification purposes
r_k21[nrow(r_k21) + 1,] = c("dummy",0,0,"U",0,0,"U")
mk21=confusionMatrix(factor(r_k21$classp),factor(r_k21$class))
mk21

r_k24=r %>% select(id,ha_k24,hb_k24,class) %>% 
  mutate(scoreA=ha_k24*sf_a[4],scoreB=hb_k24*sf_b[4], classp=if_else(scoreA > scoreB, "A",if_else(scoreA == scoreB, "U","B")))
#we add a dummy class for classification purposes
r_k24[nrow(r_k24) + 1,] = c("dummy",0,0,"U",0,0,"U")
mk24=confusionMatrix(factor(r_k24$classp),factor(r_k24$class))
mk24
```




## Random forest classifier
### model build
```{r randomf}
library(randomForest)
library(caret)
library(e1071)
library(caTools)

feats=colnames(r)[2:length(r)]
rs=r[,feats]
ca=r[r$class=="A",]
cb=r[r$class=="B",]
ca=ca[sample(nrow(cb),nrow(cb)),]
rs=rbind(ca,cb)
rs=r[,feats]

rs$class=factor(rs$class)
sample = sample.split(rs$class, SplitRatio = 0.75)
train = subset(rs, sample == TRUE) # 75% training
test  = subset(rs, sample == FALSE) # 25% for evaluation
dim(train)
dim(test)
rf <- randomForest(
  class ~ .,
  #importance=T,
  data=train,
  keep.forest = TRUE
  )
```
### model evaluation

```{r randomeval, fig.width=6, fig.height=6}
#we make the predictions
pred = predict(rf, newdata=test)
m=confusionMatrix(pred, test$class)

varImpPlot(rf,main="Importance of features")
```

we build a consolidate plot by k-mer, random forest and class

```{r code}
library(knitr)
s=mk15$byClass[,5:7]
s=rbind(s,mk18$byClass[,5:7])
s=rbind(s,mk21$byClass[,5:7])
s=rbind(s,mk24$byClass[,5:7])
s=rbind(s,m$byClass[5:7])
s=rbind(s,m$byClass[5:7])
sd=as.data.frame(s)
sd$k=c(rep(15,3),rep(18,3),rep(21,3),rep(24,3),rep("rf",2))
sd$class=c(rep(c("Paternal","Maternal","U"),4),c("Paternal","Maternal"))
kable(sd,caption = "Precision, recall and F1 for haplotype classifiers")
```

we make the plots

```{r plots, fig.width=10, fig.height=6}
library(patchwork)
p1=sd %>% filter(class != "U") %>% ggplot(aes(y=Precision, x=factor(k),fill=class)) + 
  geom_col(position="dodge2") +
  xlab("classifier: k-mer and RF")+
  theme_minimal() +
  coord_cartesian(ylim = c(0.95, 1))
p2=sd %>% filter(class != "U") %>% ggplot(aes(y=Recall, x=factor(k),fill=class)) +
  geom_col(position="dodge2")  +
  xlab("classifier: k-mer and RF")+
  theme_minimal()+ 
  coord_cartesian(ylim = c(0.9, 1))
# we merge plots
p1+p2+plot_layout(guides = 'collect') & theme(text = element_text(size = 20)) 

```


# R info

```{r}
devtools::session_info()
```
