---
title: "Tidyverse_COVID_Example"
author: "Alex Di Genova"
date: "`r Sys.Date()`"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## COVID 19 data analysis
This Rmarkdown is designed to handle COVID 19 data and produce figures as well as informative analysis.

### Demographic data of COVID19 samples

We start plotting a Chilean map
```{r map}
library(ggplot2)
library(rnaturalearth)

# Obtener los datos geográficos de Chile
chile_map <- ne_countries(country = "Chile", returnclass = "sf")

# Coordenadas aproximadas de Santiago y Rancagua
locations <- data.frame(
  city = c("Santiago", "Rancagua"),
  lon = c(-70.6483, -70.7398),
  lat = c(-33.4569, -34.1708)
)

# Crear el mapa utilizando ggplot2 y resaltar Santiago y Rancagua
p1=ggplot(data = chile_map) +
  geom_sf() +
  geom_point(data = locations, aes(x = lon, y = lat, color = city, shape=city), size = 3) +
  scale_color_manual(values = c("Santiago" = "blue", "Rancagua" = "red")) +
  ggtitle("Chile") +
  theme_classic() +
  theme(text = element_text(size = 12), legend.position = "none",  axis.text.x=element_blank()) +
  labs(color = "Ciudad")
p1
```
We get demographic data
```{r democovid}
library(tidyverse)
library(ggbeeswarm)
df=read.table("data/all_samples.txt", h=T)
df=df %>% separate(Date, into=c("m","d","y")) %>% mutate(Ct=as.integer(Ct))

p2=df %>% ggplot(aes(x=y,y=Age,color=Region,shape=Sex)) + geom_quasirandom() +
scale_color_manual(values = c("Metropolitana" = "blue", "O'Higgins" = "red")) + labs(x="Years",y="Age",title="Samples")+
  theme_minimal()
p2
```

We merge plot1 and plot2
```{r, mergedplots, fig.width=5,fig.height=3}
library(patchwork)
(p1|p2) + plot_annotation(tag_levels = 'A')
pdf("Fig1.pdf",5,3)
(p1|p2) + plot_annotation(tag_levels = 'A')
dev.off()
```
### Intra-Host Heterogeneity

We compute the number of iSNV present in the dataset and perform several statistical analyses

#### loading the data

```{r loadisnv}
csv_files <- list.files(pattern="*.variants.tsv",path = "data/SNVs/")

l_data <- list()
for (file in csv_files) {
  data <- read.table(paste0("data/SNVs/",file),h=T)
  #data <- data %>% filter(ALT_FREQ >=0.95)
  data$Filename <- file
  l_data[[file]] <- data
}
# we merge all list dataframes by row
SNV_ivar <- bind_rows(l_data)
SNV_ivar=SNV_ivar %>% separate(Filename,into=c(NA,NA,"barcode"), remove=TRUE,extra="drop") %>% mutate(barcode=as.integer(barcode))

glimpse(SNV_ivar)
```
